---
title: "Introduction"
output:
  html_document:
    highlight: textmate
    toc_float: false
    includes:
      before_body: [include_header.html, include_lessons_nav.html]
      after_body: [include_lessons_nav_end.html, include_footer.html]
---

Introduction to R language
========================================================

* Large Data Sets
* Modeling and simulations
* Statistical analysis
* Extensive computational task

-----

* Genomics, transcryptomycs
* Computational evolutionary biolo
* Structural biology
* Image analysis
* Quantative ecology

----

#### 3 diensions of languages

* __High__ vs low
* __General__ vs targeted
* __Interpretive__ vs compiled

----

#### Structure of R language

* S language
* Functions written in R itself

----

#### Specific task

* Multiple testing
* MCMC
* Bias 
* Gauss
* Plotting
* Sequencing

----
#### Check graphics system


```{r}
plot(density(rnorm(100)), col = "red")
```



----
#### Variables types


```{r}
x = 5
x <- 5
assign("x", 5)
```

----
#### Commands and Basic Functions

function c()


```{r}
x <- c(1, 3, 6, 7)
```

----
#### Vectors


* numerical
* character
* logical

(factors are not classic vector objects)


```{r}
char <- c("8", "edge", "turn8k")
logic <- c(TRUE, FALSE, TRUE, FALSE)
```


### Dynamic typisation


```{r}
dynam <- c(8, "dds", TRUE)
dynam <- c(1, 0, 1, 1, TRUE)
dynam
```

Priority Char->integer->Logic

----

#### Functions

x=function(arg1,arg2,...)

Invocation - function call

args:

* tagged(named)
* positional

Positional:
log(x,base)

```{r}
log(x, 10)
```

Tagged: 
log(base=base,x)

```{r}
log(base = 10, x)
```

----
#### Logical operations

>,<,<=,>=,==

```{r}
x = 10
x < 5
```

```{r}
x > 5
```

```{r}
x == 10
```

----

#### Math operations

+, -, *, /, ^, sin(), cos(), sqrt(), log()


```{r}
x^10
```

----

Quiz

```{r}
x = 10/(5 - 5)
x
```

----
#### Recycling Rule


```{r}
x <- c(2, 4, 5, 3, -4, 0)
x * 5
```

```{r}
y <- c(5, 3)
x * y
```

```{r}
z <- c(1, 2, 3, 4)
x * z
```

----
#### Length function

```{r}
length(x)
```

----
#### More basic functions

log(), mean(), min(), max(), sum(), prod(), abs()

```{r}
x <- c(3, 4, 6, -7, 8)
mean = sum(x)/length(x)
mean
```

----
#### Regular sequenses

```{r}
x = 2:10
y = 5:1
z = -5:10
w = -5:-10
x
```

```{r}
y
```

```{r}
z
```

```{r}
w
```

----

#### Regular sequence

seq(from=,to=,by=)


```{r}
z = seq(2, 5, 0.5)
x = seq(3, 9.1, 0.1)
y = seq(9, 3, -0.2)
length(x)
```

```{r}
length(y)
```

```{r}
x * y
```

----
#### Regular sequence

rep(x, times=)

```{r}
x = c("A", "B", "C")
rep(x, 4)
```

----
#### Logical vectors

mode(x)

```{r}
x <- 1:10
y <- x <= -5
mode(x)
```


```{r}
mode(y)
```

```{r}
mode(mode(x))
```

----
#### Vector indexation 

R is 1 based language

```{r}
x = 1:20
x[2]
```

```{r}
x[0]
```

```{r}
x[5:9]
```

```{r}
x[c(3, 7, 8, 9, 20)]
```

```{r}
x[-19]
```

```{r}
x[-(1:6)]
```

```{r}
x[x > 10]
```

```{r}
x[TRUE]
```

```{r}
y = c(TRUE, FALSE)
x[y]
```

----
#### Adding to vectors


```{r}
vector <- c()
v2 <- c(TRUE, FALSE, TRUE, TRUE, FALSE, TRUE)
vector = c(v2, vector)
vector = c(F, vector)
x = 1:3
x = c(5, x)
x
```
----
#### Named vectors

```{r}
x = 1:5
names(x)
```

```{r}
names(x) <- c("Fisrt", "Second", "Third", "Forth", 
    "First")
names(x)
```

```{r}
x["First"]
```

----

#### Vector attributes

length()
names()
mode()

----


```{r}
log(3) == log(3, base = exp(1))
```

```{r}
x = seq(5, 100, 0.1)
x[length(x)]
```

----
#### Sort and order functions


```{r}
x <- c(-10, 2, 78, 34, -98, 23, 0, -8)
order(x)
```

```{r}
sort(x)
```

```{r}
sort(x) == x[order(x)]
```





