---
title: "Itroduction to R"
author: "Alex Yaroslavtsev"
output:
  rmdshower::shower_presentation:
    self_contained: false
    katex: true
    ratio: 16x10
    css: styles.css
---

#Introduction to R language

---

* Large Data Sets
* Modeling and simulations
* Statistical analysis
* Extensive computational task

---

* Genomics, transcryptomycs
* Computational evolutionary biolo
* Structural biology
* Image analysis
* Quantative ecology

----

## 3 diensions of languages

* __High__ vs low
* __General__ vs targeted
* __Interpretive__ vs compiled

----

## Structure of R language

* S language
* Functions written in R itself

----

## Specific task

* Multiple testing
* MCMC
* Bias 
* Gauss
* Plotting
* Sequencing

----

## Check graphics system

```{r}
plot(density(rnorm(100)), col = "red")
```

----

## Variables types

```{r}
x = 5
x <- 5
assign("x", 5)
```

----

## Commands and Basic Functions

function c()


```{r}
x <- c(1, 3, 6, 7)
```

----

## Vectors


* numerical
* character
* logical

(factors are not classic vector objects)


```{r}
char <- c("8", "edge", "turn8k")
logic <- c(TRUE, FALSE, TRUE, FALSE)
```

---

### Dynamic typisation


```{r}
dynam <- c(8, "dds", TRUE)
dynam <- c(1, 0, 1, 1, TRUE)
dynam
```

Priority Char->integer->Logic

----

## Functions

x=function(arg1,arg2,...)

Invocation - function call

args:

* tagged(named)
* positional

Positional:
log(x,base)

---

## Functions

```{r}
log(x, 10)
```

Tagged: 
log(base=base,x)

```{r}
log(base = 10, x)
```

----

## Logical operations

<, >, <=, >=, ==

```{r}
x = 10
x < 5
```

----

## Logical operations

```{r}
x > 5
```

```{r}
x == 10
```

----

## Math operations

+, -, *, /, ^, sin(), cos(), sqrt(), log()

```{r}
x^10
```

----

## Quiz

```{r}
x = 10/(5 - 5)
x
```

----

## Recycling Rule

```{r}
x <- c(2, 4, 5, 3, -4, 0)
x * 5
```

```{r}
y <- c(5, 3)
x * y
```

----

## Recycling Rule

```{r}
x
z <- c(1, 2, 3, 4)
x * z
```

---

## Length function

```{r}
length(x)
```

---

## More basic functions

log(), mean(), min(), max(), sum(), prod(), abs()

```{r}
x <- c(3, 4, 6, -7, 8)
mean = sum(x)/length(x)
mean
```

----

## Regular sequenses

```{r}
x = 2:10
y = 5:1
z = -5:10
w = -5:-10
x
```

```{r}
y
```

----

## Regular sequence

```{r}
z
```

```{r}
w
```

----

## Regular sequence

seq(from=,to=,by=)


```{r}
z = seq(2, 5, 0.5)
x = seq(3, 9.1, 0.1)
y = seq(9, 3, -0.2)
length(x)
```

----

## Regular sequence

```{r}
length(y)
```

```{r}
x * y
```

----

## Regular sequence

rep(x, times=)

```{r}
x = c("A", "B", "C")
rep(x, 4)
```

----

## Logical vectors

mode(x)

```{r}
x <- 1:10
y <- x <= -5
mode(x)
```

----

## Logical vectors

```{r}
mode(y)
```

```{r}
mode(mode(x))
```

---

## Vector indexation 

R is 1 based language

```{r}
x = 1:20
x[2]
```

```{r}
x[0]
```

---

## Vector indexation 

```{r}
x[5:9]
```

```{r}
x[c(3, 7, 8, 9, 20)]
```

---

## Vector indexation 

```{r}
x[-19]
```

```{r}
x[-(1:6)]
```

---

## Vector indexation 

```{r}
x[x > 10]
```

```{r}
x[TRUE]
```

```{r}
y = c(TRUE, FALSE)
x[y]
```

----

## Adding to vectors

```{r}
vector <- c()
v2 <- c(TRUE, FALSE, TRUE, TRUE, FALSE, TRUE)
vector = c(v2, vector)
vector = c(F, vector)
x = 1:3
x = c(5, x)
x
```

----

## Named vectors

```{r}
x = 1:5
names(x)
```

```{r}
names(x) <- c("First", "Second", "Third", "Forth", "Fith")
names(x)
```

```{r}
x["First"]
```

----

## Vector attributes

length()
names()
mode()

---

## Some tricks

```{r}
log(3) == log(3, base = exp(1))
```

```{r}
x = seq(5, 100, 0.1)
x[length(x)]
```

----

## Sort and order functions


```{r}
x <- c(-10, 2, 78, 34, -98, 23, 0, -8)
order(x)
```

```{r}
sort(x)
```

```{r}
sort(x) == x[order(x)]
```

---

## Quiz 

#### Turn vector x upside down

---

## Quiz - turn vector upside down

```{r}
a <- seq(-8,10,.5)
a
x = a[length(a):1]
x
```

---

## Quiz

#### Insert in middle of vector a - 99

---

## Quiz - insert in middle of vector a - 99

```{r}
y = c(a[1:round(length(a)/2)], 99, a[(round(length(a)/2) + 1):length(a)])
```

---

## Quiz 

#### Prove arcos(cos(x)) = x

---

## Quiz prove arcos(cos(x)) = x

```{r}
x = 0:3
y = acos(cos(x))
x == y
```

---

## Lets remember that we can name vectores

```{r}
x = c(sin(10), log(5,8), log(3,exp(1)), log(38.6,10), cos(pi + 1), exp(1)^(cos(0)), atan(+Inf))
names(x) = c("sin(10)", "log8(5)", "ln3", "lg(38.6)", "cos(pi+1)", "e^cos(0)", "arctn(+Inf)");
sort(x)
```

---

## Boolean operations

* ___! = not___
* ___| = or___
* ___& = and___

A U B - conjunction

---

## Quiz - Boolean operations

```{r}
x = seq(-10,10,.01)
```

#### Get all numbers from -10 to -3 and from 0 to 6, but not 3


---

## Quiz - Boolean operations

```{r}
x = seq(-10,10,.01)
y = x[(x > -10 & x < -3) | (x > 0 & x < 6 & x != 3)]
y
```

---

## Factors and summary() function

x = factor("Mouse")
```{r}
f = c(rep("mouse", 3), rep("rat", 2), rep("fly", 11))
f = rep(c("mouse", "rat", "fly"), c(3, 2, 11))
x = factor(f)
summary(x)
summary(f)
```

---

## Factors and summary() function

```{r}
summary(1:10)
summary(1:10)[4]
```

---

## Quiz - factors

#### Make vector of DNA and calculate ratio of each nucleotide
```{r}
dna=factor(rep(c("T","G","C","A"),c(16,38,6,10)))
```
---

## Quiz - factors

```{r}

dna=factor(rep(c("T","G","C","A"),c(16,38,6,10)))
ratio=summary(dna)/sum(summary(dna))
ratio
```

---

## Function sample()

sample:

* mixing
* generating

---

## Function sample()

```{r}
dna
sample(dna)

#sample(dna,size = 84, replace = FALSE) - will give error
```
---

## Function sample()

```{r}

sample(dna,size = 84, replace = TRUE)
sampledna = sample(c("A","T","G","C"), size = 10000, replace = TRUE)
ratio = summary(factor(sampledna)) / length(sampledna)
ratio2 = summary(factor(dna)) / length(dna)
```

---

## Function sample()

```{r}
ratio
ratio2
```

---

## Functions union(x,y),intersect(x,y),setdiff(x,y)

```{r}
x=1:10
y=6:14
union(x,y)
intersect(x,y)
setdiff(x,y)
setdiff(y,x)
```

---

## Removing all variables in session

```{r}
rm(list=ls())
```

---

## Function cumsum()

```{r}
cumsum(1:10)
```

---

## Quiz 
#### make big sequence with ratio of G=0.5

---

## Quiz make big sequence with ratio of G=0.5

```{r}
dna2=sample(c("A","T","G","G","G","C"), size=3826513,replace=TRUE)
ratio=summary(factor(dna2))/length(dna2)
ratio
dna2=sample(c("A","T","G","C"), size=3826513,prob=c(.5/3,.5/3,.5,.5/3),replace=TRUE)
ratio=summary(factor(dna2))/length(dna2)
ratio

```

---

## Subsettings

```{r}
iris[iris$Species==c("setosa","virginica"),]
iris[iris$Species=="setosa" | iris$Species=="virginica",c("Petal.Length","Species")]

```

---

## All main data types

Type      | Function    | Options
----------|-------------|-------------------------------------
Vector    | c()         | [i] -position or logic
Factor    | factor()    | [] - position or logic
Matrix    | matrix()    | [row,col]; [i] - positional logical
List      | list()      | [[i]], $
dataframe | data.frame()| [,] $

---

## Quiz - subsetting

#### Find species with Petal.Length>2  but <3, show them and their abundance

---

## Quiz - subsetting

```{r}
summary(factor(iris[(iris$Petal.Length>2 & iris$Petal.Length<6),"Species" ]))
```

---

#### Functions

```{r}
my_func = function(p) 
  {
    x=p^2
    y=x^3
    print("result")
    return(c(x,y))
  }
my_func(2)
```

---

#### Functions

```{r}
mathses = function(p,q)
  {
    result = c(p+q,log(p+q,10),exp(p+q))
    names(result) = c("summ", "log","exp")
    return(result)
  }
mathses(2,3)
```

---

## Conditional instructions if functions

```{r}

mf = function(p)
  {
  
    if (p>0)
      {
        x=log(p)
        return(x)
      }
    else
      {
        return(NULL)      
      }
  }
mf(-1)
```

---

##Quiz conditional questions

### gamma(x) gives !(x-1), we need to make function which makes true factorial

---

##Quiz conditional questions

```{r}
myfactorial = function(p) 
  {
    if (p>0 & p%%1==0)
      {return(gamma(p+1))}
    else
      {return(NULL)}
  }
  
  myfactorial(5)
```
  
---

##Quiz conditional questions

```{r}
 myfactorial = function(p) 
  {
    if (p>0 & p%%1==0)
      {return(gamma(p+1))}
    else
      {return(NULL)}
  }

```


